Redundant calibration as a complex optimization problem

### 2017-01-06 paper.tex
structure improved even more

The most standard way of solving Eq.~\eqref{eq:least_squares} is to use a gradient type of minimization algorithm. In this paper we will be focusing on the following gradient minimization algorithms: Gauss-Newton and Levenberg-Marquardt \citep{Levenberg1944,Marquardt1963}. 
The GN update step is defined as
\begin{equation}
 \Delta \bzeta = (\bJ^T\bJ)^{-1} \bJ^T\brho,
\end{equation}
while the LM update takes the following form
\begin{equation}
 \Delta \bzeta = (\bJ^T\bJ+\lambda\bD)^{-1} \bJ^T\brho,
\end{equation}
where $\bJ$ is the Jacobian matrix (i.e. the derivative of the model $\bnu$ with respect to the parameter vector $\bzeta$), $\lambda$ denotes the 
damping factor, while $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. In this paper we denote 
matrix transposition with $(*)^T$.
These update steps can be used in an iterative manner to determine new estimates of $\bzeta$.

@@

\section{Problem Statement}
In Section~\ref{sec:s_mes}, we present the scalar measurement equation. Next, we define some indexing functions in Section~\ref{sec:indexing}. The measurement 
equation and the indexing functions from Section~\ref{sec:indexing} are then used in Section~\ref{sec:r_mes} to phrase redundant calibration as a least-squares problem, i.e. Eq.~\eqref{eq:least_squares_red}.
We present the Signal-to-noise ratio definition we used in this paper in Section\ref{sec:snr}. The general scope and limitations of this paper is discussed in Section~\ref{sec:scope}.

\subsection{Scalar Measurement Equation}
\label{sec:s_mes}
The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ is described by
\begin{equation}
\label{eq:vis_definition}
d_{pq} = g_{p}\conj{g_q}y_{pq} + n_{pq},
\end{equation}
where $g_{p}$ and $g_{q}$ are the direction independent gains associated with antenna $p$ and $q$, $y_{pq}$ denotes the true visibility that baseline $pq$ measured
and $n_{pq}$ is the thermal noise component. Eq.~\eqref{eq:vis_definition} is known as the scalar measurement equation. Conjugation is denoted by $\conj{(*)}$. The true value of $g_p$, $g_q$ and $y_{pq}$ are unknown. We try to estimate the value of
$g_p$, $g_q$ and $y_{pq}$ by calibrating our data and then imaging the result.

@@

Recall that redundant baselines sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. 
The aforementioned fact allows us to group all the redundant baselines in a redundant array into the same redundant baseline group. When our array 
is a redundant array we therefore only require single index values for each redundant baseline group, instead of having a unique single index label for 
each and every baseline. Let $\phi_{pq}$ be the function that maps composite antenna pairs to the appropriate single redundant baseline group index. Note that $\phi_{pq}$ is not symmetric as 
$\phi_{pq} = 0$ if $p>q$. We can also define the following symmetric variant of $\phi_{pq}$:
\begin{equation}
\zeta_{pq} = 
\begin{cases}
\phi_{pq}~\textrm{if}~p \leq q\\
\phi_{qp}~\textrm{if}~p>q
\end{cases}.
\end{equation} 
It is possible to construct a simple analytic expression for $\zeta_{pq}$ when our array is in an east-west regular configuration, namely $\zeta_{pq} = |q-p|$. 
It becomes increasingly more difficult to construct analytic expressions of $\zeta_{pq}$ for other more complicated array layouts. 
We will refer to $\zeta_{pq}$ as the symmetric geometric function. The empirically constructed symmetric geometry
functions of three different redundant layouts are depicted in Fig.~\ref{fig:geometry_function}. We denote the range of $\zeta_{pq}$ with $\mathcal{R}(\zeta_{pq})$. The maximal element 
that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be construed as the maximal number of unique redundant baseline groups which can be formed for a given 
array layout. Alternatively, we can interpret the symmetric geometry functions in Fig.~\ref{fig:geometry_function} as geometry matrices. In this alternative paradigm, $\zeta_{pq}$ denotes a matrix entry instead. The dimension of these geometry 
matrices are $N\times N$.

@@

We have now defined enough notation to enable us to formulate redundant calibration as a least-squares problem:
\begin{equation}
\label{eq:least_squares_red}
\min_{\bz} \Lambda(\bz) = \min_{\bz} \|\br\|_F^2 = \min_{\bz} \|\bd - \bv(\bz)\|_F^2, 
\end{equation}
where
\begin{align}
 \by &= [y_1,\cdots,y_L]^T, & \bg &=[g_1,\cdots,g_N]^T,\nonumber\\
 \bz &= [\bg^T,\by^T]^T. &  &\label{eq:parm_definitions}
 \end{align}
The number of model parameters that we need to solve for in Eq.~\eqref{eq:least_squares_red} is equal to $P = 2(N+L)$, since redundant calibration is a complex problem.
We plot $L$ and $P$ for two different geometric layouts in Fig~\ref{fig:pl}.
The analytic expressions for the red and the blue curve in Fig.~\ref{fig:ln} are respectively equal to
\begin{equation}
L = 2N-\frac{1}{2}\sqrt{12N-3}-\frac{1}{2}, 
\end{equation}
and
\begin{equation}
L = N - 1. 
\end{equation}

@@

\subsection{Signal-to-noise Ratio}
\label{sec:snr}
The following SNR (signal-to-noise ratio) definition is used in this paper  
\begin{equation}
\label{}
\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}}, 
\end{equation}
where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baselines and $\odot$ denotes the Hadamard product. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.

>>>>

The most standard way of solving Eq.~\eqref{eq:least_squares} is to use a gradient type of minimization algorithm. In this paper we will be focusing on the following gradient minimization algorithms: Gauss-Newton and Levenberg-Marquardt \citep{Levenberg1944,Marquardt1963}. 
The GN update step is defined as
\begin{equation}
 \Delta \bzeta = (\bJ^T\bJ)^{-1} \bJ^T\brho,
\end{equation}
while the LM update takes the following form
\begin{equation}
 \Delta \bzeta = (\bJ^T\bJ+\lambda\bD)^{-1} \bJ^T\brho,
\end{equation}
where $\bJ$ is the Jacobian matrix (i.e. the derivative of the model $\bnu$ with respect to the parameter vector $\bzeta$), $\lambda$ denotes the 
damping factor, while $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. In this paper we respectively denote 
matrix transposition and matrix inversion with $(*)^T$ and $(*)^{-1}$.
These update steps can be used in an iterative manner to determine new estimates of $\bzeta$.

@@

\section{Problem Statement}
In Section~\ref{sec:s_mes}, we present the scalar measurement equation. Next, we define some indexing functions in Section~\ref{sec:indexing}. The measurement 
equation and the indexing functions from Section~\ref{sec:indexing} are then used in Section~\ref{sec:r_mes} to phrase redundant calibration as a least-squares problem, i.e. Eq.~\eqref{eq:least_squares_red}.
We present the Signal-to-noise ratio (SNR) definition we used in this paper in Section\ref{sec:snr}. The general scope and limitations of this paper is discussed in Section~\ref{sec:scope}.

\subsection{Scalar Measurement Equation}
\label{sec:s_mes}
The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ is described by
\begin{equation}
\label{eq:vis_definition}
d_{pq} = g_{p}\conj{g_q}y_{pq} + n_{pq},
\end{equation}
where $g_{p}$ and $g_{q}$ are the direction independent gains associated with antenna $p$ and $q$, $y_{pq}$ denotes the true visibility that baseline $pq$ measured
and $n_{pq}$ is the thermal noise component. Eq.~\eqref{eq:vis_definition} is known as the scalar measurement equation. Conjugation is denoted by $\conj{(*)}$. The true value of $g_p$, $g_q$ and $y_{pq}$ are unknown. We normally try to estimate the true value of
$g_p$, $g_q$ and $y_{pq}$ by calibrating our data and then imaging the result.

@@

Recall that redundant baselines sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. 
The aforementioned fact allows us to group all the redundant baselines in a redundant array into the same redundant baseline group. When our array 
is a redundant array we therefore only require single index values for each redundant baseline group, instead of a unique single index label for 
each baseline. Let $\phi_{pq}$ be the function that maps composite antenna pairs to the appropriate single redundant baseline group index. Note that $\phi_{pq}$ is not symmetric as 
$\phi_{pq} = 0$ if $p>q$. We can also define the following symmetric variant of $\phi_{pq}$:
\begin{equation}
\zeta_{pq} = 
\begin{cases}
\phi_{pq}~\textrm{if}~p \leq q\\
\phi_{qp}~\textrm{if}~p>q
\end{cases}.
\end{equation} 
It is possible to construct a simple analytic expression for $\zeta_{pq}$ when our array is in an east-west regular configuration, namely $\zeta_{pq} = |q-p|$. 
It becomes increasingly more difficult to construct analytic expressions of $\zeta_{pq}$ for other more complicated array layouts. 
We will refer to $\zeta_{pq}$ as the symmetric geometric function. The empirically constructed symmetric geometry
functions of three different redundant layouts are depicted in Fig.~\ref{fig:geometry_function}. We denote the range of $\zeta_{pq}$ with $\mathcal{R}(\zeta_{pq})$. The maximal element 
that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be construed as the maximal number of unique redundant baseline groups which can be formed for a given 
array layout. Alternatively, we can interpret the symmetric geometry functions in Fig.~\ref{fig:geometry_function} as geometry matrices. In this alternative paradigm, $\zeta_{pq}$ denotes a matrix entry instead. The dimension of these geometry 
matrices are $N\times N$.

@@

We have now defined enough notation to enable us to formulate redundant calibration as a least-squares problem:
\begin{equation}
\label{eq:least_squares_red}
\min_{\bz} \Lambda(\bz) = \min_{\bz} \|\br\|_F^2 = \min_{\bz} \|\bd - \bv(\bz)\|_F^2, 
\end{equation}
where
\begin{align}
 \bg &=[g_1,\cdots,g_N]^T, & \by &= [y_1,\cdots,y_L]^T,\nonumber\\
 \bz &= [\bg^T,\by^T]^T. &  &\label{eq:parm_definitions}
 \end{align}
The number of model parameters that we need to solve for in Eq.~\eqref{eq:least_squares_red} is equal to $P = 2(N+L)$, since redundant calibration is a complex problem.
We plot $L$ and $P$ for two different geometric layouts in Fig~\ref{fig:pl}.
The analytic expressions for the red and the blue curve in Fig.~\ref{fig:ln} are respectively equal to
\begin{equation}
L = 2N-\frac{1}{2}\sqrt{12N-3}-\frac{1}{2}, 
\end{equation}
and
\begin{equation}
L = N - 1. 
\end{equation}

@@

\subsection{Signal-to-noise Ratio}
\label{sec:snr}
The following SNR definition is used in this paper  
\begin{equation}
\label{}
\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}}, 
\end{equation}
where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baselines and $\odot$ denotes the Hadamard product. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.
